---
title: "Object Oriented Field Guide"
author: "Ian Flores"
output: github_document
---

> DISCLAIMER: This is going to be more of a worksheet/live coding session than a presentation by itself. 

> DISCLAIMER: Most of the code here is taken directly from the Advanced R book or is a modification of code inside that book. The book is available at this link: http://adv-r.had.co.nz/OO-essentials.html

# Object-Oriented Programming

## Let's get started with an example

* Think of a blueprint of a building 
    * How about a specific building?
        * What does this building does?
        * What does this building have?

Can you think of another hierarchy example?

## How does this fit into OO Programming
* Think of a blueprint of a building - **Class**
    * How about a specific building? - **Object**
        * What does this building does? - **Methods**
        * What does this building have? - **Attributes**


# Object-Oriented Systems in R

R has (3) three OO systems:

* S3
* S4
* Reference Classes


## Base System

* Every R object is a C `struct` 
* Base types are not an object system
* To determine an object's base type, use the `typeof` function. 

```{r}
typeof(1)
```

```{r}
typeof('x')
```


* Functions are of type closure.

```{r}
f <- function() {}
typeof(f)
```

## S3

The majority of objects used in R are S3 objects. How can we know if the object we are using is in the S3 system?

```{r}
library(pryr)
df <- data.frame(num = 1:10, letter = letters[1:10])
otype(df)
```

```{r}
tib <- tibble::as.tibble(df)
otype(tib)
```

However in S3, the methods belong to the functions, not to the objects. 

```{r}
## You do
dplyr::filter(tib, num == 1)
```

```{r, eval=FALSE}
## You DON'T do
tib.filter(num == 1)
```

### Function Generics

When you call a function such as `mean` you are calling a generic which will then match the object to the corresponding method.

If you want to see all the methods a function has, you can use:

```{r}
methods('mean')
```

What if you have an object and you want to see which functions you can apply to it?

```{r}
methods(class = 'data.frame')
```

### Creating our own objects

```{r}
nums <- structure(list(x = 1:10), class = 'custom_num')
class(nums)
```

How do I test if an object is of a certain class?
```{r}
inherits(nums, 'custom_num')
```

However, if using already created class is better to use the default way as converting between classes does not always work. 
```{r, eval = FALSE}
as.data.frame(nums)
```

```{r}
data.frame(x = 1:10)
```

You can also assign methods for a specific class the following way:
```{r}
print.custom_num = function(x){print(x$x)}
print.custom_num(nums)
```

Let's see if we can create a new data_scientist object, with class `Person` with at least 2 attributes. 
```{r}
## Your Code Here
data_scientist <- NULL
if (inherits(data_scientist, 'Person')){
    print("Success!")
}
```

## S4 

S4 objects are not that common in the R ecosystem. However, it is a rich OO system for development in R. 

How to identify S4 objects?

```{r}
library(stats4)
y <- c(1:10)
nLL <- function(lambda) - sum(dpois(y, lambda, log = TRUE))
fit <- mle(nLL, start = list(lambda = 5), nobs = length(y))
```

```{r}
## From the base package
isS4(fit)
```

How to identify from which class inherits an object?
```{r}
is(fit)
```

```{r}
is(fit, 'mle')
```

### Creating our own S4 objects

S4 objects have 3 key properties:
* A name
* A named list of slots
* A string giving the class inheritance

```{r}
setClass('r_programmer', slots = list(name = 'character', packages = 'character'))
```

```{r}
(ian <- new('r_programmer', name = 'Ian', packages = c('ggplot2', 'leaflet')))
```

```{r}
## Create your own S4 class and an object. 
```

How do we access the package `leaflet`? 
```{r}
ian
```

```{r}
ian@packages
```

```{r}
ian@packages[2]
```

#### Brief parentheses about inheritance

If an S4 object contains an S3 class or a base type, it will have a `.Data` slot. 
```{r}
setClass('RangedNumeric', contains = 'numeric', slots = list(min = 'numeric', max = 'numeric'))
```

```{r}
(rn <- new('RangedNumeric', 1:5, min = 1, max = 5))
```

```{r}
rn@.Data
```

S4 comes with a `standardGeneric()` function to assign the generic call of a function. 

Hadley Wickham added some reources in his book to go more in depth into the S4 system. 

## Reference Classes

** Very similar to classes in `Python` **

If we want to create a Reference class we use the `setRefClass()` function

```{r}
vehicle <- setRefClass("vehicle")
vehicle$new()
```

How about if we want to add  attributes?
```{r}
(vehicle <- setRefClass("vehicle", 
                       fields = list(tires = "numeric", doors = "numeric")))
```

#### Let's create some vehicle objects
```{r}
bike <- vehicle$new(tires = 2, doors = 0)
car <- vehicle$new(tires = 4, doors = 4)
```

What if I want to create a copy of bike?
```{r}
bike_copy <- bike$copy()
```

However, if you modify the original object, the copy will not be affected as well. 

```{r}
bike$tires <- 3
bike
```

```{r}
bike_copy
```

As compared to S3 and S4 methods are part of objects, just as in Python or C++.

```{r}
vehicle <- setRefClass("Vehicle", 
                       fields = list(tires = 'numeric', doors = 'numeric'),
                       methods = list(
                           steal_all_tires = function(x) {
                               tires <<- 0
                           },
                           seal_one_door = function(x) {
                               doors <<- doors - 1
                           }, 
                           convert_vehicle = function(x){
                               tires <<- tires + 1
                           }
                       ))

```

```{r}
bike <- vehicle$new(tires = 2, doors = 0)
bike
```

```{r}
bike$convert_vehicle()
bike
```

How can I inherit from a previous class?

```{r}
buses <- setRefClass("Buses",
                     contains = "Vehicle",
                     fields = list(people_onboard = 'numeric'),
                     methods = list(
                         add_people = function(x){
                             people_onboard <<- people_onboard + x
                         }
                             )
                    )
```

```{r}
bus_480 <- buses$new(tires = 8, doors = 2, people_onboard = 25)
bus_480
```

How can we test if `bus_480` is a Reference Class?

```{r}
(is(bus_480, "refClass"))
```

```{r}
otype(bus_480)
```